---
title: "Class 2-2, Part 1"
author: "Health Data Analysis Practicum (AS.280.347)"
date: "February 21, 2024"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = TRUE)
knitr::opts_knit$set(root.dir = "/cloud/project")
```


```{r libraries, include = FALSE}
library(tidyverse)  
library(haven)
library(broom)
library(survey)
library(ggpubr)
library(ggrepel)
library(kableExtra)
library(plotrix)
library(knitr)  

```

## Initial data inspection (continued)

```{r read-data}
dat <- read_sas('module_2/data/d.sas7bdat')
rename <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           race = DMQ_14_1,
           gender = GENDER,
           born = US_BORN,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           bmi = BMI,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = EXAM_WT)
summary(rename)
```


### Non-categorical variables 

There are four non-categorial variables that we will use in our analysis:

  * `id`: Sample case ID, unique to each individual in the sample
  * `age`: Sample age, range `r min(rename$age)`-`r max(rename$age)` years
  * `bmi`: BMI = $kg/m^2$ where $kg$ is a person's weight in kilograms and $m$ is their height in meters
  * `surveyweight`: Numeric values associated with each observation to let us know how much weight the observation should receive in our analysis (more details later)
  
### Categorical variables 

We will consider ten categorical variables. Note that the levels of these variables are detailed in the [Variable Codebook](https://med.nyu.edu/departments-institutes/population-health/divisions-sections-centers/epidemiology/sites/default/files/nyc-hanes-datasets-and-resources-public-dataset-codebook.pdf){target="_blank"}. 

  * `race`: 
    + 100 = White
    + 110 = Black/African American
    + 120 = Indian
    + 140 = Native Hawaiian/Other Pacific Islander
    + 180 = Asian
    + 250 = Other race
  * `gender`:
    + 1 = Male
    + 2 = Female
  * `born`:
    + 1 = US born
    + 2 = Other country
  * `diet`: 
    + 1 = Excellent
    + 2 = Very good 
    + 3 = Good
    + 4 = Fair
    + 5 = Poor
  * `diabetes`: Has person ever been told by a doctor or health professional that they have diabetes or sugar diabetes?
    + 1 = Yes
    + 2 = No 
    + 3 = Prediabetes
  * `cholesterol`: Has person ever been told by a doctor or health professional that their blood cholesterol was high?
    + 1 = Yes
    + 2 = No
  * `drink`: In the past 12 months, how often did sample drink any type of alcoholic beverage
    + 1 = Weekly
    + 2 = Monthly
    + 3 = Yearly
  * `smoke`: 
    + 1 = Never smoker
    + 2 = Current smoker
    + 3 = Former smoker
  * `income`:
    + 1 = Less than $20,000
    + 2 = $20,000 - $39,999
    + 3 = $40,000 - $59,999
    + 4 = $60,000 - $79,999
    + 5 = $80,000 - $99,999
    + 6 = $100,000 or more
  * `hypertension`: Has person ever been told by a doctor or health professional that they have hypertension or high blood pressure?
    + 1 = Yes
    + 2 = No
    


## Adjust data types

From the data summaries above, we can see that there are several 
categorical variables like `race`, `gender`, `born`, `diet`, `income`, `diabetes`, `bmi`, `drink`, and `smoke`, which are currently being 
treated as numerical values (giving means/medians in the summary) but are really categorical variables that should be treated as factors. This is because the categories are simply labeled with numbers.

We want to convert these categorical variables to factors using the numerical values and category labels given in the 
[Variable Codebook](https://med.nyu.edu/departments-institutes/population-health/divisions-sections-centers/epidemiology/sites/default/files/nyc-hanes-datasets-and-resources-public-dataset-codebook.pdf){target="_blank"} 
and shown earlier in this document.

We can use the `factor()` function in base R to convert each variable and assign the correct levels. Any values that are not included in the `levels` argument will get set to `NA` values. We also want to think about creating a natural ordering to the factor levels here: the first level will generally be our reference level in a regression model, so it makes sense to try to give them an order that reflects our choice of reference group.  For example, we will probably want to examine diet in increasing order of how good it is, so we order the levels from 5 to 1, rather than 1 to 5, and assign the labels appropriately.

```{r recodeFactors}
hy_df <- rename %>% mutate(race=factor(race, levels=c(100, 110, 120, 140, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Indian /Alaska Native', 
                              'Pacific Islander', 
                              'Asian', 'Other Race')),
                     gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     born = factor(born, levels = c(1,2),
                        labels = c("US Born", "Non-US Born")),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,1,3), 
                          labels=c('No','Yes','Prediabetes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(1,2,3), 
                       labels=c('Weekly', 'Monthly', 'Yearly')),
                     smoking = factor(smoking, levels=c(3:1), 
                         labels=c('Never smoker','Former smoker','Current smoker')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes'))
                     )

```

If we look at a summary of our data set again, we see that now the `summary()` function gives us counts for each category of these variables rather than the meaningless numerical summaries, like means, it was giving before.
```{r}
summary(hy_df)
```


## Follow up on questions from last class

### Coding of hypertension variable

Notice the ordering of the hypertension variable above. This will make no hypertension the baseline, making it easier to interpret any modeling output with hypertension as the outcome.

### Drink coding/missing values

Using the [Variable Codebook](https://med.nyu.edu/departments-institutes/population-health/divisions-sections-centers/epidemiology/sites/default/files/nyc-hanes-datasets-and-resources-public-dataset-codebook.pdf){target="_blank"}, we see that the variable we've chosen for `drink`, `ALQ_1_UNIT` in the original survey data, seems to be a follow-up question to the earlier variable `ALQ_1`, which asks how often did the survey participant drink any type of alcoholic beverage.  For this first question, a value of 0 means they never drink.

Let's look at the frequency of counts of this first `ALQ_1` variable with the function `count()`.

```{r}
dat %>% count(ALQ_1) %>% print(n=Inf)
```

There are `r sum(dat$ALQ_1 == 0, na.rm = TRUE)` people who indicated that they never drink.  These individuals would not have answered the following question (`AlQ_1_unit`) to give a frequency of drinking in the past 12 months.  Now we see why there are so many missing values for `drink`. Among these `r sum(is.na(hy_df$drink))` missing values, `r sum(dat$ALQ_1 == 0, na.rm = TRUE)` samples never drink and there are just `r sum(is.na(hy_df$drink)) - sum(dat$ALQ_1 == 0, na.rm = TRUE)` actual missing values. Therefore, merging these two variables as one is a better way to capture drinking that includes those who never drink. 

Discuss the following questions in your groups:

(1) Write some code to create a new `drink` variable which has the value 4 whenever someone answered `0` to variable `ALQ_1`, and the usual values of 1, 2 or 3 for the `drink` variable otherwise.

(2) In our analysis, we will probably want to compare individuals who drink weekly, monthly, or yearly to those who never drink. What do you think is the best order of the levels `Never`, `Weekly`, `Monthly` and `Yearly`?

(3) Find one or two other variables in the code chunk above where you could argue that the levels of the factors should be reordered.



