---
title: "Class 2-8: Wrapping up Module 2"
author: "Health Data Analysis Practicum (AS.280.347)"
date: "March 13, 2024"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = TRUE)
knitr::opts_knit$set(root.dir = "/cloud/project")
```

## Module 2: Factors that are associated with development of hypertension

Recall that our main questions of interest are:

  * Question 2.1: What factors measured in the NYC HANES survey are associated with having hypertension?
  * Question 2.2: How do our estimates from survey-weighted logistic regression differ from those where we ignore survey weights?


The data science learning objectives for this module include:

  * Understand the components of a data analysis report
  * Gain experience performing data cleaning, and assessing whether you have been successful
  * Practice selecting data visualizations that fit into the context of your statistical analysis

The statistical learning objectives for this module include:

  * Gain further experience with logistic regression and selecting an appropriate model for your question
  * Understand what a survey-weighted analysis is and how/when we perform one
  * Learn how to select survey weights for unbalanced data


## Reminder: What are the data?

For this case study, we will use data from the [New York City (NYC) Health and Nutrition Examination Survey (NYC HANES)](http://nychanes.org/){target="_blank"}, modeled on the [National Health and Nutrition Examination Survey (NHANES)](https://wwwn.cdc.gov/nchs/nhanes/default.aspx){target="_blank"}. NHANES is a population-based, cross-sectional study with data collected from a physical examination and laboratory tests, as well as a face-to-face interview and an audio computer-assisted self-interview (ACASI). It is designed to assess the health and nutritional status of adults and children in the United States. NYC HANES is a local version of NHANES, which implies it mainly focuses on the New York area. 


## Learning objectives for this week

Our main question of interest for this module is: Based on the data collected from NYC HANES, which risk factors play a role in development of hypertension?

Today, we will continue to work toward answering this by learning how to:

* Describe what makes an effective comparison of survey-weighted and unweighted model results and share some examples of tables and figures that some of you included with your assignments this week.
* Touch base on any questions about model selection and selecting the correct survey weights to use in your analysis.
* Understand how to include inline R code and why it is useful.
* Discuss specific best practices for writing statistical results




## A few general comments as you work toward Module 2 final project

* In general, more numeric values should be included, both in your summaries of your data visualizations and in your interpretations of your model results. We will talk more about this today!
* Remember that best practices would be to include only those variables that you will use in your model in the final data set that you include in your analysis. This is in order to reduce the number of observations lost due to missing data. Once you have decided on your final model, I would suggest going back to the beginning and adjusting your code to achieve this. This is also a good time to double check that you are using the appropriate variable for your survey weighting (which most of you are doing at this point).
* Avoid language about hypothesis testing; just state the result and give the p-value.
* Remember to clean things up -- remove any extra text or code, make all code and unformatted R output invisible
* Be sure to reference and describe what you find from each table or figure you include in the report
* Variable names in displays and tables should be nicely "human readable" -- no concatenations of the variable name and level
* Your model results should not include the intercept and should be on the odds ratio (exponentiated) scale
* For comparing the weighted and unweighted results, some indication of agreement or disagreement in statistical significance should be included, either with a p-value or a CI
* When you present results for categorical variables, whether in a visualization or a model output, be sure it is clear to the reader what the baseline level is.


## Example of reordering factor levels

```{r, echo = FALSE}
library(tidyverse)  
library(knitr)  
library(ggpubr)
library(ggrepel)
library(tidyverse)
library(kableExtra)
library(survey)
library(haven)
library(broom)
library(plotrix)
library(pander)
library(gridExtra)

dat <- read_sas('./module_2/data/d.sas7bdat')
dim(dat)
```

```{r, echo=FALSE}

rename <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           ageCat= AGEGRP5C,
           race = DMQ_14_1,
           gender = GENDER,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = CAPI_WT)

# combine the two alcohol variables into one and combine 'Indian /Alaska Native' and 'Pacific Islander' categories in the race variable into "Other Race" because their sample sizes are too small 
rename <- rename %>% 
          mutate(drink = ifelse(dat$ALQ_1 == 0, 4, drink),
                 race = ifelse(race == 120 | race == 140, 250, race))

# recode race, income, drink, smoking, and hypertension variables 
hy_df <- rename %>% mutate(race=factor(race, levels=c(100, 110, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Asian', 'Other Race')),
                      gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,3, 1), 
                          labels=c('No','Prediabetes','Yes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(4:1), 
                       labels=c('Never','Yearly', 'Monthly', 'Weekly')),
                     smoking = factor(smoking, levels=c(1,3,2), 
                                            labels=c('Never smoker','Former smoker','Current smoker')),
                     ageCat = factor(ageCat, levels=c(1:5), 
                                     labels=c('20-29', '30-39', '40-49', '50-59', '>=60')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes'))
                     )
# filter out NA
hy_p_df <- 
  hy_df %>%
  drop_na()


hypertension_design <- svydesign(
  ids = ~1,
  weights = ~hy_df$surveyweight,
  data = hy_df
)

#survey-weighted summary statistics
h_design_nona <- subset(hypertension_design, complete.cases(hy_df))

g1 <- svyglm(hypertension ~ 
               age + race + income + 
               diabetes + cholesterol,
             family = binomial(link = 'logit'), 
             design = h_design_nona)

g1_res <- tidy(g1, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95) %>% 
  filter(term != "(Intercept)")  
### unweighted logistic regression
g1_noWt <- glm(hypertension ~ 
               age + race + income + 
               diabetes + cholesterol,
              family = quasibinomial(link = 'logit'), data=hy_p_df)

g1_noWt_res <- tidy(g1_noWt, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95) %>% 
  filter(term != "(Intercept)") 


#combine these two results into one data frame
g1_res <- g1_res %>% mutate(method="Weighted")
g1_noWt_res <- g1_noWt_res %>% mutate(method="Unweighted")
g1_both<-rbind(g1_res, g1_noWt_res)

#rename estimate as odds ratio
names(g1_both)[names(g1_both) == "estimate"] <- "odds.ratio"


#data visualizations comparing the results separating categorical and continuous variables 
g1_both %>% 
  ggplot(aes(x = odds.ratio, y = term), group_by(method)) +
     geom_vline(xintercept = 1, linetype = "dashed") +
     geom_errorbarh(aes(y=term, xmin = conf.low, 
     xmax = conf.high, color = method), linewidth = 0.8, position="dodge")+
     geom_point(aes(color = method,
                    shape = method), size = 3, position=position_dodge(width=0.9))+
  labs(title = "hypertension weighted vs unweighted logistic regression", 
       subtitle = "odds ratio with 95% confidence interval for categorical terms") + 
  xlab("odds ratio") +
  ylab("categorical terms")


#data visualizations comparing the results separating categorical and continuous variables 
g1_both %>% 
  ggplot(aes(x = odds.ratio, y = fct_reorder(term, odds.ratio))) +
     geom_vline(xintercept = 1, linetype = "dashed") +
     geom_errorbarh(aes(xmin = conf.low, 
     xmax = conf.high, color = method), linewidth = 0.8, position="dodge")+
     geom_point(aes(color = method,
                    shape = method), size = 3, position=position_dodge(width=0.9))+
  labs(title = "hypertension weighted vs unweighted logistic regression", 
       subtitle = "odds ratio with 95% confidence interval for categorical terms") + 
  xlab("odds ratio") +
  ylab("categorical terms")



```

How do we improve the term labels while making sure the order is what we want it to be?

```{r}
g1_both <- g1_both %>% mutate(term = sub("age", "Age", term), 
                              term = sub("race", "Race: ", term),
                              term = sub("income", "Income: ", term),
                              term = sub("diabetes", "Diabetes: ", term),
                              term = sub("cholesterol", "Cholesterol: ", term))

#data visualizations comparing the results separating categorical and continuous variables 
g1_both %>% 
  ggplot(aes(x = odds.ratio, y = factor(term, levels = term, labels = term))) +
     geom_vline(xintercept = 1, linetype = "dashed") +
     geom_errorbarh(aes(xmin = conf.low, 
     xmax = conf.high, color = method), linewidth = 0.8, position="dodge")+
     geom_point(aes(color = method,
                    shape = method), size = 3, position=position_dodge(width=0.9))+
  labs(title = "hypertension weighted vs unweighted logistic regression", 
       subtitle = "odds ratio with 95% confidence interval for categorical terms") + 
  xlab("odds ratio") +
  ylab("categorical terms")


```


## Scientific writing for your reports

Read the Methods and Results section of one of the two JAMA articles included in the class materials repository today. What do you observe about what is contained in each section? What can we say about the use of numeric support and the way statistical significance is treated?


## Assignment 2.4 (final project for Module 2)

Write a complete report investigating the questions of interest from this module. Your report should include:

* Some simple/concise background on the data set and a clear presentation of the question you want to answer;
* An exploratory data analysis, where you use data visualizations (tables and figures) to illustrate relationships between variables in the data set, as they relate to the question you are interested in answering, and then discuss what these visualizations tell you;
* Model results presented in a tidy way, perhaps with a nice table of relevant coefficients and p-values, including a clear discussion of the effects of the variables included in the model;
* A comparison of results from the survey-weighted model to those from a standard glm, and a brief discussion of what you see.


In the written portions of your report, think about connecting each table or visualization that you make with your downstream modeling choices. How do your displays point you toward the model you end up fitting? 

Also, think about justifying each choice of variable in your model: can you use a visualization to do this? Look at significance of coefficients? Talk about how you are curious about a specific variable and how it influences hypertension? Compare models using AIC to decide on your final model?

What is an effective way of illustrating how the results of the weighted and unweighted approaches compare?

Finally, try to make your report as readable as possible: hide any R code or warning messages, try to make the tables readable, etc. 

* Submit your updated analysis report in R Markdown through Github by Monday March 25, 2024 at midnight.
* Post a screenshot of some aspect of your final project that you think is particularly interesting on Piazza in the "Assignment 2-4 Results" thread.  Add a sentence or two about what you can conclude based on your analysis. You are welcome to post this anonymously to your classmates. You can also include comments about what your chose to do or questions you had as you were making the display and fitting your model.
* You may work together on this assignment, but you must submit your own report; please credit in your assignment anyone with whom you collaborated.
* There is no class next week due to spring break; in class two weeks from now we will continue with discussion of your final project ideas.

## Looking ahead to your final projects


It is time to start thinking about your final projects, which will be the focus of the course after we wrap up Module 2. 


### Project structure

Your final project will have the following pieces:

* Question of interest
* Data set and design
  * Outcome variable
  * Predictor variable(s) of primary interest
  * Potential confounders
  * Potention effect modifiers
* Directed acyclic graph (DAG) showing the hypothesized relationships of interest and including potential confounders and effect modifiers
* A primary analysis to address the question of interest
* Communication of results in tables and figures

### Working project document

To help you get started, we have created a working Google Doc where you can record your plans. This will help you organize your thoughts, find classmates working on similar questions, and help us keep track of who is working on what.

Here is link to the [working doc](https://docs.google.com/document/d/1KxpJ0wodHlcG814QaoSFhVpfvlvilIwPXhOBuNoV8i0/edit?usp=sharing){target="_blank"}.

Feel free to start contributing your ideas; nothing here is set in stone, this is just a starting point that you will update in the coming weeks.


### Getting started

Shortly after we return from Spring Break, we will want you to have identified:

* A research question of interest in public health
* A data source you can use to answer this question

It will be helpful if you spend some time thinking about this now! 

**Note:** This is often an iterative process where you decide on a question, look for data, and then realize you need to refine or modify your question based on available data.

Here are some tips for **framing a research question in public health**:

* Start with a *general* area of public health in which you have interest; then narrow to a *specific* frame your question in terms of investigating a relationship between a specific outcome variable (like "disease status" from our Module 1) and one or more primary predictor variables ("smoking status" from our Module 1.)
* Later you will need to think about the possibility of effect modifiers and possible confounders, but for now just think about that primary relationship of interest.

Here are some tips for **locating data to answer this question**:

* If you have a specific area of interest in mind, you can Google for data in that area.
* Or explore the links below to see what type of data is available.
* Collections of publicly-available datasets via U.S. government, WHO, World Bank, and other public and private entities:
  * https://www.healthdata.gov/browse?limitTo=datasets
  * https://guides.lib.berkeley.edu/publichealth/healthstatistics/rawdata
  * http://www.datasciencecentral.com/profiles/blogs/10-great-healthcare-data-sets
  * https://www.cdc.gov/nchs/data_access/ftp_data.htm
  * https://data.cdc.gov/browse
  * https://www.icpsr.umich.edu/web/pages/
  * https://data.worldbank.org/
  * https://data.gov/
  * https://www.kaggle.com/
  * https://whscldatablog.wordpress.com/category/public-access-data/
  * https://www.who.int/data/collections
  * https://data.census.gov/cedsci/table?q=United%20States 

* Study-specific datasets: 
  * Longitudinal Study of American Youth: https://www.icpsr.umich.edu/web/ICPSR/studies/30263?q=LSAY
  * Demographic and Health Survey (DHS): https://dhsprogram.com/data/Using-DataSets-for-Analysis.cfm 
  * National Health Interview Survey (NHIS): https://www.cdc.gov/nchs/nhis/index.htm
  * National Health and Nutrition Examination Survey (NHANES): https://wwwn.cdc.gov/nchs/nhanes/Default.aspx
  * National Comorbidity Survey (NCS): https://www.icpsr.umich.edu/web/ICPSR/studies/06693
  * NYC HANES: https://www1.nyc.gov/site/doh/data/data-sets/nyc-hanes-info.page
  * NYC Youth Risk Behavior Study: https://www1.nyc.gov/site/doh/data/data-sets/nyc-youth-risk-behavior-survey.page

