---
title: "Class 2-7: Wrapping up Module 2"
author: "Health Data Analysis Practicum (AS.280.347)"
date: "March 11, 2024"
output: 
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = TRUE)
knitr::opts_knit$set(root.dir = "/cloud/project")
```

## Module 2: Factors that are associated with development of hypertension

Recall that our main questions of interest are:

  * Question 2.1: What factors measured in the NYC HANES survey are associated with having hypertension?
  * Question 2.2: How do our estimates from survey-weighted logistic regression differ from those where we ignore survey weights?


The data science learning objectives for this module include:

  * Understand the components of a data analysis report
  * Gain experience performing data cleaning, and assessing whether you have been successful
  * Practice selecting data visualizations that fit into the context of your statistical analysis

The statistical learning objectives for this module include:

  * Gain further experience with logistic regression and selecting an appropriate model for your question
  * Understand what a survey-weighted analysis is and how/when we perform one
  * Learn how to select survey weights for unbalanced data


## Reminder: What are the data?

For this case study, we will use data from the [New York City (NYC) Health and Nutrition Examination Survey (NYC HANES)](http://nychanes.org/){target="_blank"}, modeled on the [National Health and Nutrition Examination Survey (NHANES)](https://wwwn.cdc.gov/nchs/nhanes/default.aspx){target="_blank"}. NHANES is a population-based, cross-sectional study with data collected from a physical examination and laboratory tests, as well as a face-to-face interview and an audio computer-assisted self-interview (ACASI). It is designed to assess the health and nutritional status of adults and children in the United States. NYC HANES is a local version of NHANES, which implies it mainly focuses on the New York area. 


## Learning objectives for this week

Our main question of interest for this module is: Based on the data collected from NYC HANES, which risk factors play a role in development of hypertension?

Today, we will continue to work toward answering this by learning how to:

* Describe what makes an effective comparison of survey-weighted and unweighted model results and share some examples of tables and figures to help inspire you to improve your own visualizations
* Touch base on any questions about model selection and selecting the correct survey weights to use in your analysis.
* Review how to include inline R code and why it is useful.
* Discuss specific best practices for writing statistical results



Start by loading libraries and raw data set.
```{r}
library(tidyverse)  
library(knitr)  
library(ggpubr)
library(ggrepel)
library(tidyverse)
library(kableExtra)
library(survey)
library(haven)
library(broom)
library(plotrix)
library(pander)
library(gridExtra)

dat <- read_sas('./module_2/data/d.sas7bdat')
dim(dat)
```

```{r , echo=FALSE}
hy_df <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           race = DMQ_14_1,
           gender = GENDER,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           bmi = BMI,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = EXAM_WT)

hy_df <- hy_df %>% 
          mutate(drink = ifelse(dat$ALQ_1 == 0, 4, drink))

hy_df <- hy_df %>% mutate(race=factor(race, levels=c(100, 110, 120, 140, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Indian /Alaska Native', 
                              'Pacific Islander', 
                              'Asian', 'Other Race')),
                     gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,3, 1), 
                          labels=c('No','Prediabetes','Yes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(4:1), 
                       labels=c('Never','Yearly', 'Monthly', 'Weekly')),
                     smoking = factor(smoking, levels=c(1,3,2), 
                         labels=c('Never smoker','Former smoker','Current smoker')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes')),
                     )

## we will not use this in our survey design object, but will use it for visualizations below
hy_p_df <-
  hy_df %>%
  drop_na()


hypertension_design <- svydesign(
  id = ~1,
  weights = ~hy_df$surveyweight,
  data = hy_df
)
h_design_nona <- subset(hypertension_design, complete.cases(hy_df))
```



## Getting to Question 2.2

Recall Question 2.2: How do our estimates from survey-weighted logistic regression differ from those where we ignore survey weights?

To answer this, use your final model from your survey-weighted analysis and fit a standard (unweighted) logistic regression instead. How might you want to compare the results of these two models? A table? Some kind of visualization?

Some of you posted examples of tables and figures in your work this week on Piazza. We will start with some group discussion of the figures you created as well as some additional examples from previous years.


### Example plots

In your groups, for each plot, discuss things that you like and things that you think could be improved. 


#### Example 1


```{r , echo = FALSE}

rename <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           race = DMQ_14_1,
           gender = GENDER,
           born = US_BORN,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           bmi = BMI,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = EXAM_WT)

rename <- rename %>% 
          mutate(drink = ifelse(dat$ALQ_1 == 0, 4, drink))

hy_df <- rename %>% mutate(race=factor(race, levels=c(100, 110, 120, 140, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Indian /Alaska Native', 
                              'Pacific Islander', 
                              'Asian', 'Other Race')),
                     gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     born = factor(born, levels = c(1,2),
                        labels = c("US Born", "Non-US Born")),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,1,3), 
                          labels=c('No','Yes','Prediabetes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(4:1), 
                       labels=c('Never','Yearly', 'Monthly', 'Weekly')),
                     smoking = factor(smoking, levels=c(1,3,2), 
                         labels=c('Never smoker','Former smoker','Current smoker')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes'))
                     )

hy_p_df <- 
  hy_df %>%
  drop_na()

hypertension_design <- svydesign(
  ids = ~1,
  weights = ~hy_df$surveyweight,
  data = hy_df
)

h_design_nona <- subset(hypertension_design, complete.cases(hy_df))

g <- svyglm(hypertension ~ age + gender + bmi + cholesterol + diabetes + race + income, 
    family = quasibinomial(link = 'logit'), design = h_design_nona)


g_results <- tidy(g, exponentiate = TRUE) %>%
  filter(term != "(Intercept)")

g_results$term <- c("Age", "Female", "BMI", "High Cholesterol", "Diabetes", "Prediabetes", "Black/African American", "Indian/Alaska Native", "Pacific Islander", "Asian", "Other Race", "Income $20,000 - $39,999", "Income $40,000 - $59,999", "Income $60,000 - $79,999", "Income $80,000 - $99,999", "Income $100,000 or more")

g1_noWt <- glm(hypertension ~ 
               age + gender + bmi + cholesterol + diabetes + race + income,
              family = quasibinomial(link = 'logit'), data=hy_p_df)

g1_noWt_res <- tidy(g1_noWt, exponentiate = TRUE) %>% 
  filter(term != "(Intercept)")

g1_noWt_res$term <- c("Age", "Female", "BMI", "High Cholesterol", "Diabetes", "Prediabetes", "Black/African American", "Indian/Alaska Native", "Pacific Islander", "Asian", "Other Race", "Income $20,000 - $39,999", "Income $40,000 - $59,999", "Income $60,000 - $79,999", "Income $80,000 - $99,999", "Income $100,000 or more")

g_results <- g_results %>% mutate(method="Weighted")
g1_noWt_res <- g1_noWt_res %>% mutate(method="Unweighted")

g1_both<-rbind(g_results, g1_noWt_res)

g1_both %>% ggplot(aes(x = estimate, y = term),
     group_by(g1_both$method)) +
     geom_point(aes(color = factor(method),
                    shape = factor(method)),
                size = 3) +
    labs(title = "Comparison of weighted and unweighted analysis",
       x = "Odds Ratio",
       y = "Variable") +
   theme_minimal()


```

#### Example 2


```{r, echo=FALSE}

rename <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           ageCat= AGEGRP5C,
           race = DMQ_14_1,
           gender = GENDER,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = CAPI_WT)

# combine the two alcohol variables into one and combine 'Indian /Alaska Native' and 'Pacific Islander' categories in the race variable into "Other Race" because their sample sizes are too small 
rename <- rename %>% 
          mutate(drink = ifelse(dat$ALQ_1 == 0, 4, drink),
                 race = ifelse(race == 120 | race == 140, 250, race))

# recode race, income, drink, smoking, and hypertension variables 
hy_df <- rename %>% mutate(race=factor(race, levels=c(100, 110, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Asian', 'Other Race')),
                      gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,3, 1), 
                          labels=c('No','Prediabetes','Yes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(4:1), 
                       labels=c('Never','Yearly', 'Monthly', 'Weekly')),
                     smoking = factor(smoking, levels=c(1,3,2), 
                                            labels=c('Never smoker','Former smoker','Current smoker')),
                     ageCat = factor(ageCat, levels=c(1:5), 
                                     labels=c('20-29', '30-39', '40-49', '50-59', '>=60')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes'))
                     )
# filter out NA
hy_p_df <- 
  hy_df %>%
  drop_na()


hypertension_design <- svydesign(
  ids = ~1,
  weights = ~hy_df$surveyweight,
  data = hy_df
)

#survey-weighted summary statistics
h_design_nona <- subset(hypertension_design, complete.cases(hy_df))

g1 <- svyglm(hypertension ~ 
               age + race + income + 
               diabetes + cholesterol,
             family = binomial(link = 'logit'), 
             design = h_design_nona)

g1_res <- tidy(g1, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95) %>% 
  filter(term != "(Intercept)")  
### unweighted logistic regression
g1_noWt <- glm(hypertension ~ 
               age + race + income + 
               diabetes + cholesterol,
              family = quasibinomial(link = 'logit'), data=hy_p_df)

g1_noWt_res <- tidy(g1_noWt, exponentiate = TRUE, conf.int = TRUE, conf.level = 0.95) %>% 
  filter(term != "(Intercept)") 


#combine these two results into one data frame
g1_res <- g1_res %>% mutate(method="Weighted")
g1_noWt_res <- g1_noWt_res %>% mutate(method="Unweighted")
g1_both<-rbind(g1_res, g1_noWt_res)

#rename estimate as odds ratio
names(g1_both)[names(g1_both) == "estimate"] <- "odds.ratio"

#data visualizations comparing the results separating categorical and continuous variables 
categorical <- g1_both %>% 
  filter(term != "age") %>% 
  ggplot(aes(x = odds.ratio, y = term),
     group_by(g1_both$method)) +
     geom_vline(xintercept = 0, linetype = "dashed") +
     geom_errorbarh(aes(y=term, xmin = conf.low, 
     xmax = conf.high, color = method), linewidth = 0.8, position="dodge")+
     geom_point(aes(color = method,
                    shape = method), size = 3, position=position_dodge(width=0.9))+
  labs(title = "hypertension weighted vs unweighted logistic regression", 
       subtitle = "odds ratio with 95% confidence interval for categorical terms") + 
  xlab("odds ratio") +
  ylab("categorical terms")

continuous <- g1_both %>%
  filter(term == "age") %>% 
  ggplot(aes(x = odds.ratio, y = term),
     group_by(g1_both$method)) +
     geom_errorbarh(aes(y=term, xmin = conf.low, 
     xmax = conf.high, color = method), linewidth = 0.8, position="dodge")+
     geom_point(aes(color = method,
                    shape = method), size = 3, position=position_dodge(width=0.9))+
  labs(title = "hypertension weighted vs unweighted logistic regression", 
       subtitle = "odds ratio with 95% confidence interval for continuous terms") + 
  xlab("odds ratio") +
  ylab("continuous terms")

ggarrange(categorical, continuous, ncol=1, nrow=2, heights=c(2,1))

```

#### Example 3

```{r, echo = FALSE}

hy_df <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           race = DMQ_14_1,
           gender = GENDER,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           bmi = BMI,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = EXAM_WT)

hy_df <- hy_df %>% 
          mutate(drink = ifelse(dat$ALQ_1 == 0, 4, drink))

hy_df <- hy_df %>% mutate(race=factor(race, levels=c(100, 110, 120, 140, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Indian /Alaska Native', 
                              'Pacific Islander', 
                              'Asian', 'Other Race')),
                     gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,3, 1), 
                          labels=c('No','Prediabetes','Yes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(4:1), 
                       labels=c('Never','Yearly', 'Monthly', 'Weekly')),
                     smoking = factor(smoking, levels=c(1,3,2), 
                         labels=c('Never smoker','Former smoker','Current smoker')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes')),
                     )

## we will not use this in our survey design object, but will use it for visualizations below
hy_p_df <-
  hy_df %>%
  drop_na()


hypertension_design <- svydesign(
  id = ~1,
  weights = ~hy_df$surveyweight,
  data = hy_df
)
h_design_nona <- subset(hypertension_design, complete.cases(hy_df))

#weighted logistic regression
g1 <- svyglm(hypertension ~ 
              gender + diabetes + cholesterol + smoking,
              family = quasibinomial(link = 'logit'), 
              design = hypertension_design)

g1_res <- tidy(g1)
g1_res <- cbind(g1_res, confint(g1))

### unweighted logistic regression
g1_noWt <- glm(hypertension ~ 
               gender + diabetes + cholesterol + smoking,
               family = quasibinomial(link = 'logit'), data=hy_p_df)
g1_noWt_res <- tidy(g1_noWt)
g1_noWt_res <- cbind(g1_noWt_res, confint(g1_noWt))

### Combine results
g1_res <- g1_res %>% mutate(method="Weighted")
g1_noWt_res <- g1_noWt_res %>% mutate(method="Unweighted")
g1_both <- rbind(g1_res, g1_noWt_res)

### Plot results
g1_both %>% 
  ggplot(aes(x = estimate, y = term)) +
     geom_vline(xintercept = 0, linetype = "dashed") +
     geom_errorbarh(aes(estimate, term, xmin = `2.5 %`, xmax = `97.5 %`, color = factor(method)), height = 0.2, size = 0.8) +
     geom_point(aes(color = factor(method),
                    shape = factor(method)),size = 3) +
  facet_wrap(~method, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("#0072B2", "#D55E00")) +
  scale_shape_manual (values = c(19, 17)) +
  labs(title = expression(bold("95% confidence interval")),
       subtitle = "Weighted vs. unweighted logistic regression", x = "Odds ratio", y = "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12, face = "italic"), 
        legend.position = "bottom",
        axis.title.y = element_blank(),
        panel.spacing = unit(0.5, "lines"),
        strip.text = element_text(size = 12, face = "bold"))

```


#### Example 4

How do you think the variables on this plot are ordered? 

```{r, echo= FALSE}
dat <- read_sas('./module_2/data/d.sas7bdat')
rename <- 
  dat %>% 
    select(id = KEY,
           age = SPAGE,
           race = DMQ_14_1,
           gender = GENDER,
           born = US_BORN,
           diet = DBQ_1,
           income = INC20K,
           diabetes = DIQ_1,
           bmi = BMI,
           cholesterol = BPQ_16,
           drink = ALQ_1_UNIT,
           smoking = SMOKER3CAT,
           hypertension = BPQ_2,
           surveyweight = CAPI_WT)

rename <- rename %>% 
          mutate(drink = ifelse(dat$ALQ_1 == 0, 4, drink))

hy_df <- rename %>% mutate(race=factor(race, levels=c(100, 110, 120, 140, 180, 250), 
                      labels=c('White', 'Black/African American', 
                              'Indian /Alaska Native', 
                              'Pacific Islander', 
                              'Asian', 'Other Race')),
                     gender = factor(gender, levels=c(1,2), 
                        labels=c('Male', 'Female')),
                     born = factor(born, levels = c(1,2),
                        labels = c("US Born", "Non-US Born")),
                     diet = factor(diet, levels=c(5:1), 
                      labels=c('Poor', 'Fair', 'Good', 
                               'Very good','Excellent')),
                     income = factor(income, levels=c(1:6), 
                        labels=c('Less than $20,000','$20,000 - $39,999',
                                 '$40,000 - $59,999','$60,000 - $79,999',
                                 '$80,000 - $99,999','$100,000 or more')),
                     diabetes = factor(diabetes, levels=c(2,3, 1), 
                          labels=c('No','Prediabetes','Yes')),
                     cholesterol = factor(cholesterol, levels=c(2,1), 
                             labels=c('Low value','High value')),
                     drink = factor(drink, levels=c(4:1), 
                       labels=c('Never','Yearly', 'Monthly', 'Weekly')),
                     smoking = factor(smoking, levels=c(1,3,2), 
                                      labels=c('Never smoker','Former smoker','Current smoker')),
                     hypertension = factor(hypertension, levels=c(2,1), 
                              labels=c('No','Yes'))
                     )
hy_p_df <- 
  hy_df %>%
  drop_na()



hypertension_design <- svydesign(
  ids = ~1,
  weights = ~hy_df$surveyweight,
  data = hy_df
)

h_design_nona <- subset(hypertension_design, complete.cases(hy_df))

g1 <- svyglm(hypertension ~ 
               age + gender + diet + diabetes + cholesterol,
             family = quasibinomial(link = 'logit'), 
             design = h_design_nona)

g1_results <- tidy(g1,
                exponentiate = TRUE,
                conf.int = TRUE)

g1_res <- tidy(g1, 
                conf.int = TRUE)

g1_results$term <- c("Intercept", 
                    "Age", 
                    "Gender", 
                    "Fair Diet", 
                    "Good Diet", 
                    "Very Good Diet", 
                    "Excellent Diet", 
                    "Pre-Diabetes", 
                    "Diabetes", 
                    "High Cholesterol")

g1_res$term <- c("Intercept", 
                  "Age", 
                  "Gender", 
                  "Fair Diet", 
                  "Good Diet", 
                  "Very Good Diet", 
                  "Excellent Diet", 
                  "Pre-Diabetes", 
                  "Diabetes", 
                  "High Cholesterol")

g1_noWt <- glm(hypertension ~ 
               age + gender + diet + diabetes + cholesterol,
              family = quasibinomial(link = 'logit'), data=hy_p_df)

g1_noWt_results <- tidy(g1_noWt,
                        exponentiate = TRUE,
                        conf.int = TRUE)

g1_noWt_res <- tidy(g1_noWt, 
                    conf.int = TRUE)

g1_noWt_results$term <- c("Intercept", 
                      "Age", 
                      "Gender", 
                      "Fair Diet", 
                      "Good Diet", 
                      "Very Good Diet", 
                      "Excellent Diet", 
                      "Pre-Diabetes", 
                      "Diabetes", 
                      "High Cholesterol")

g1_noWt_res$term <- c("Intercept", 
                      "Age", 
                      "Gender", 
                      "Fair Diet", 
                      "Good Diet", 
                      "Very Good Diet", 
                      "Excellent Diet", 
                      "Pre-Diabetes", 
                      "Diabetes", 
                      "High Cholesterol")

g1_res <- g1_results %>% mutate(Method="Weighted")
g1_noWt_res <- g1_noWt_results %>% mutate(Method="Unweighted")
g_both <- rbind(g1_res, g1_noWt_res)

comparedgraph <- g_both %>%
  filter(term != "Intercept") %>%
  ggplot(aes(x = estimate, y = fct_reorder(term, estimate), color = Method)) +
  geom_vline(xintercept = 1,
             linetype = "dashed") +
  geom_errorbarh(aes(xmin = conf.low, 
                     xmax = conf.high,
                     width = 0.3),
                 height = 0.5, 
                 size = 0.8,) +
  geom_point(aes(color = factor(Method), size = 3) +
  scale_color_manual(values = c("#0072B2", "#f08080"))) +
  scale_shape_manual(values = c(19, 17)) +
  theme_minimal() +
  labs(title = "Comparison of Survey-Weighted and Non-Survery Weighted Regression Analysis",
       y = "Variable",
       x = "Odds Ratio Estimate",
       caption = "Figure 4: Odds Ratio Estimates of Survey-Weighted and Non-Survery-Weighted Regression Analysis"
      )

comparedgraph 

```


#### Example 5

This example uses a function from the `jtools` package, which you can read more about [here](https://jtools.jacob-long.com){target="_blank"}. We also need to install the `ggstance` and `broom.mixed` packages to get this to work without an error. This function is powerful, but does sweep some things under the rug, so you might want to be careful using it. Once you have discussed how to improve the plot, see if you can figure out an additional argument to pass to the function to improve it.

```{r , echo = FALSE}
library(jtools)
plot_summs(g1, g1_noWt, scale = TRUE, model.names = c("Survey-Weighted", "Non Survey-Weighted"))
```




### An alternative: example tables

We know how to create nice tables of the model outputs from each method on its own.

```{r }
### survey-weighted logistic regression
g1 <- svyglm(hypertension ~ 
               bmi + age + cholesterol + income + 
               diabetes,
              family = quasibinomial(link = 'logit'), 
             design = h_design_nona)

g1_res <- tidy(g1, exponentiate = TRUE, conf.int = TRUE)
```


```{r}
pander(g1_res)
```

We then want to fit the same model using un-weighted logistic regression: 

```{r}
### unweighted logistic regression
g1_noWt <- glm(hypertension ~ 
               bmi + age + cholesterol + income + 
               diabetes,
              family = quasibinomial(link = 'logit'), data=hy_p_df)

g1_noWt_res <- tidy(g1_noWt, exponentiate = TRUE, conf.int = TRUE)
pander(g1_noWt_res)
```


Now that we have both versions of the results, we can work on combining them in different ways to make some tables.

A couple of you posted an example of a table comparison on Piazza [here](https://piazza.com/class/lrh8j0ur7id7kt/post/42_f2){target="_blank"}, [here] (https://piazza.com/class/lrh8j0ur7id7kt/post/42_f8){target="_blank"}  and [here](https://piazza.com/class/lrh8j0ur7id7kt/post/42_f6){target="_blank"}. What are some things that work well or could be improved with this set of tables?

To show the results in the same table, we want to start by joining the results from the two tables.
```{r}
g1_join <- inner_join(g1_res, g1_noWt_res, by = "term", suffix = c(".Wt", ".noWt"))
```


This example includes some nice grouped labels for the weighted and unweighted results, using the `add_header_above` function from `kableExtra`:

```{r}
tblcap.3 <- "Table 3: Comparison of weighted and unweighted results"
comp_tbl <- g1_join %>%
  select(term, `OR` = `estimate.Wt`,`Std Error` = `std.error.Wt`, `p-value` = `p.value.Wt`, `OR ` = `estimate.noWt`, `Std Error ` = `std.error.noWt`, `p-value ` = `p.value.noWt`) %>% 
  kable(format = "html",
        digits = 2,
        caption = tblcap.3) %>%
  add_header_above(c(" ","Survey Weighted" = 3, "Non-weighted" = 3))%>%
  kable_styling(full_width = TRUE)

comp_tbl

```

One thing I would like to improve about this table is the way the p-values are displayed as 0's. There are several approaches to changing this; I am showing one here using the `formatC` function.  You can see some further examples of how to play around with this [here](https://stackoverflow.com/questions/39623636/forcing-r-output-to-be-scientific-notation-with-at-most-two-decimals){target="_blank"}.


```{r}
tblcap.3 <- "Table 3: Comparison of weighted and unweighted results"
comp_tbl <- g1_join %>%
  mutate(p.value.Wt = formatC(p.value.Wt, digits = 2, format = "g"), p.value.noWt = formatC(p.value.noWt, digits = 2, format = "g")) %>% 
  select(term, `OR` = `estimate.Wt`,`Std Error` = `std.error.Wt`, `p-value` = `p.value.Wt`, `OR ` = `estimate.noWt`, `Std Error ` = `std.error.noWt`, `p-value ` = `p.value.noWt`) %>% 
  kable(format = "html",
        digits = 2,
        caption = tblcap.3) %>%
  add_header_above(c(" ","Survey Weighted" = 3, "Non-weighted" = 3))%>%
  kable_styling(full_width = TRUE)

comp_tbl

```

Here is a similar example that also includes a difference in the odds ratios:

```{r , echo = FALSE}

g_join <- inner_join(g1_results, g1_noWt_results, by = "term", suffix = c(".Wt", ".noWt"))

combinetable <- g_join %>%
  filter(term != "Intercept") %>%
  mutate(conf.Wt = paste0("(", round(conf.low.Wt, 2), ", ", round(conf.high.Wt, 2), ")")) %>%
  mutate(conf.noWt = paste0("(", round(conf.low.noWt, 2), ", ", round(conf.high.noWt, 2), ")")) %>%
  mutate(p.valueSN.Wt = formatC(p.value.Wt, digits = 2, format = "e")) %>%
  mutate(p.valueSN.noWt = formatC(p.value.noWt, digits = 2, format = "e")) %>%
  select(Term = term,
         `Odds Ratio` = `estimate.Wt`,
         `P-Value` = `p.valueSN.Wt`,
         `95% Confidence Interval` = conf.Wt,
         `Odds Ratio ` = `estimate.noWt`,
         `P-Value ` = `p.valueSN.noWt`,
         `95% Confidence Interval ` = conf.noWt) %>%
  mutate(Difference = abs(`Odds Ratio` - `Odds Ratio `)) %>%
  kable(format = "html",
        digits = 3,
        caption = "Table 3: Comparison of Weighted and Unweighted Modeling Results",) %>%
  add_header_above(c("", "Survery Weighted" = 3, "Non-Weighted" = 3, " ")) %>%
  kable_styling(full_width = TRUE)

combinetable

```
 


There is another package that makes really nice looking tables, the package `gtsummary`, with the function `tbl_regression`. It is a little fidgety with the variable names for some reason, but the tables also include the baseline levels of the variables, for example, which is quite useful. There is a tutorial about how to use this package [here](https://cran.r-project.org/web/packages/gtsummary/vignettes/tbl_regression.html){target="_blank"}.

```{r}
library(gtsummary)

tbl1<-tbl_regression(g1, exponentiate = TRUE, label = list(age = "age", bmi = "bmi"))%>%
  bold_labels()%>%
  bold_p()


tbl2<-tbl_regression(g1_noWt, exponentiate=TRUE, label = list(age = "age", bmi = "bmi"))%>%
  bold_labels()%>%
  bold_p()

tbl_merged <-
  tbl_merge(
    tbls = list(tbl1, tbl2),
    tab_spanner = c("**Survey Weighted**", "**Unweighted**")
  )

tbl_merged

```



## Model selection and survey weighting

We have talked about a couple approaches to model selection, including looking at statistical significance of variables and using the AIC. As we are combining our survey-weighting with model comparisons, it is a little tricky, since the weights we use depend on what variables are included in the model. And to use the AIC, we need to be using the same data set, i.e., including the same weights and data observations for any data sets we are comparing. This may require some fiddling to get it right.

Remember that the weight variable that you use will depend on the set of variables that will be included in your final model. You may need to revisit this choice depending on what your final model is.

Are there any questions/comments/concerns about model selection and survey-weighting?


## What is inline R code and why is it useful?



```{r}
### survey-weighted logistic regression -- not including age
g1 <- svyglm(hypertension ~ 
               bmi + cholesterol + income + 
               diabetes,
              family = quasibinomial(link = 'logit'), 
             design = h_design_nona)

g1_res <- tidy(g1, exponentiate = TRUE, conf.int = TRUE)
pander(g1_res)
```

Suppose I wanted to write some sentences to summarize what this model output is telling us. I could run the code, look at the numbers, and then type them here:

For example:
Holding all other variables constant, a one-unit increase in BMI is associated with a 7.8% increase in the odds of hypertension (p-value 2.93x10-9, 95% CI 5.2-10.5%). 

Instead of running the code and copying the numbers from the output, you could pull them directly from the R objects using code:
Holding all other variables constant, a one-unit increase in age is associated with  `r round(100*(g1_res %>% filter(term == "bmi") %>% pull(estimate)) - 100,1)`% increase in the odds of hypertension (p-value `r format(g1_res %>% filter(term=="bmi") %>% pull(p.value), digits=3)`, 95% CI `r round(100*(g1_res %>% filter(term=="bmi") %>% pull(conf.low))-100,1)`-`r round(100*(g1_res %>% filter(term=="bmi") %>% pull(conf.high))-100,1)`%). 

Although you can't see it in the knitted document, this second paragraph does not contain the numbers themselves, but R code to generate and format the numbers that should be included in the model write-up. For example, to generate the percent increase, the code is: `round(100*(g1_res %>% filter(term == "bmi") %>% pull(estimate)) - 100,1)`. And to generate the p-value, it is `format(g1_res %>% filter(term=="bmi") %>% pull(p.value), digits=3)`.


Why is this useful? 

* It is not susceptible to typos, i.e., copying and pasting the wrong numbers.
* It is easy to update, for example, if you fit a new model. Suppose you realize that you actually meant to include age as a covariate. 

```{r}
### survey-weighted logistic regression -- not including bmi
g1 <- svyglm(hypertension ~ 
               age + bmi + cholesterol + income + 
               diabetes,
              family = quasibinomial(link = 'logit'), 
             design = h_design_nona)

g1_res <- tidy(g1, exponentiate = TRUE, conf.int = TRUE)
pander(g1_res)


```

If I had hard-coded numbers into my output summary, I would have to go back and re-write and double check the numbers. But if I have used code, the same code will just give me updated results:

Holding all other variables constant, a one-unit increase in age is associated with  `r round(100*(g1_res %>% filter(term == "bmi") %>% pull(estimate)) - 100,1)`% increase in the odds of hypertension (p-value `r format(g1_res %>% filter(term=="bmi") %>% pull(p.value), digits=3)`, 95% CI `r round(100*(g1_res %>% filter(term=="bmi") %>% pull(conf.low))-100,1)`-`r round(100*(g1_res %>% filter(term=="bmi") %>% pull(conf.high))-100,1)`%). 

A few tips to make things easier:

* You will probably want to save your model output and/or some processed versions in R objects that you can easily refer to in your inline code, like `g1_res` above.
* You should try to use variable names and the tidyverse piping commands as much as possible, rather than indexing into specific rows or columns of your output. For example, if you had written code referring to `age` as the third variable in the model (for example using `g1_res[3, 2]` to access the odds ratio), this would be incorrect in the new model, where `age` is included before `bmi`.

You can create whatever output you want in your R code and refer back to it using inline code:
```{r}
foo <- "this is a test"
```

Print `r foo`.

## Scientific writing for your reports

Read the Methods and Results section of one of the two JAMA articles included in the class materials repository today. What do you observe about what is contained in each section? What can we say about the use of numeric support and the way statistical significance is treated?


## Assignment 2.4 (final project for Module 2)

Write a complete report investigating the questions of interest from this module. Your report should include:

* Some simple/concise background on the data set and a clear presentation of the question you want to answer;
* An exploratory data analysis, where you use data visualizations (tables and figures) to illustrate relationships between variables in the data set, as they relate to the question you are interested in answering, and then discuss what these visualizations tell you;
* Model results presented in a tidy way, perhaps with a nice table of relevant coefficients and p-values, including a clear discussion of the effects of the variables included in the model;
* A comparison of results from the survey-weighted model to those from a standard glm, and a brief discussion of what you see.


In the written portions of your report, think about connecting each table or visualization that you make with your downstream modeling choices. How do your displays point you toward the model you end up fitting? 

Also, think about justifying each choice of variable in your model: can you use a visualization to do this? Look at significance of coefficients? Talk about how you are curious about a specific variable and how it influences hypertension? Compare models using AIC to decide on your final model?

What is an effective way of illustrating how the results of the weighted and unweighted approaches compare?

Finally, try to make your report as readable as possible: hide any R code or warning messages, try to make the tables readable, etc. 

* Submit your updated analysis report in R Markdown through Github by Monday March 27, 2023 at midnight.
* Post a screenshot of some aspect of your final project that you think is particularly interesting on Piazza in the "Assignment 2-4 Results" thread.  Add a sentence or two about what you can conclude based on your analysis. You are welcome to post this anonymously to your classmates. You can also include comments about what your chose to do or questions you had as you were making the display and fitting your model.
* You may work together on this assignment, but you must submit your own report; please credit in your assignment anyone with whom you collaborated.
* There is no class next week due to spring break; in class two weeks from now we will continue with discussion of your final project ideas.

## Looking ahead to your final projects




It is time to start thinking about your final projects, which will be the focus of the course after we wrap up Module 2. 


### Project structure

Your final project will have the following pieces:

* Question of interest
* Data set and design
  * Outcome variable
  * Predictor variable(s) of primary interest
  * Potential confounders
  * Potention effect modifiers
* Directed acyclic graph (DAG) showing the hypothesized relationships of interest and including potential confounders and effect modifiers
* A primary analysis to address the question of interest
* Communication of results in tables and figures

### Working project document

To help you get started, we have created a working Google Doc where you can record your plans. This will help you organize your thoughts, find classmates working on similar questions, and help us keep track of who is working on what.

Here is link to the [working doc](https://docs.google.com/document/d/1R-8157ozdOPi-09h3EsconPV0GK9ZhxuY6hRQ5QiMpM/edit?usp=sharing){target="_blank"}.

Feel free to start contributing your ideas; nothing here is set in stone, this is just a starting point that you will update in the coming weeks.


### Getting started

Shortly after we return from Spring Break, we will want you to have identified:

* A research question of interest in public health
* A data source you can use to answer this question

It will be helpful if you spend some time thinking about this now! 

**Note:** This is often an iterative process where you decide on a question, look for data, and then realize you need to refine or modify your question based on available data.

Here are some tips for **framing a research question in public health**:

* Start with a *general* area of public health in which you have interest; then narrow to a *specific* frame your question in terms of investigating a relationship between a specific outcome variable (like "disease status" from our Module 1) and one or more primary predictor variables ("smoking status" from our Module 1.)
* Later you will need to think about the possibility of effect modifiers and possible confounders, but for now just think about that primary relationship of interest.

Here are some tips for **locating data to answer this question**:

* If you have a specific area of interest in mind, you can Google for data in that area.
* Or explore the links below to see what type of data is available.
* Collections of publicly-available datasets via U.S. government, WHO, World Bank, and other public and private entities:
  * https://www.healthdata.gov/browse?limitTo=datasets
  * https://guides.lib.berkeley.edu/publichealth/healthstatistics/rawdata
  * http://www.datasciencecentral.com/profiles/blogs/10-great-healthcare-data-sets
  * https://www.cdc.gov/nchs/data_access/ftp_data.htm
  * https://data.cdc.gov/browse
  * https://www.icpsr.umich.edu/web/pages/
  * https://data.worldbank.org/
  * https://data.gov/
  * https://www.kaggle.com/
  * https://whscldatablog.wordpress.com/category/public-access-data/
  * https://www.who.int/data/collections
  * https://data.census.gov/cedsci/table?q=United%20States 

* Study-specific datasets: 
  * Longitudinal Study of American Youth: https://www.icpsr.umich.edu/web/ICPSR/studies/30263?q=LSAY
  * Demographic and Health Survey (DHS): https://dhsprogram.com/data/Using-DataSets-for-Analysis.cfm 
  * National Health Interview Survey (NHIS): https://www.cdc.gov/nchs/nhis/index.htm
  * National Health and Nutrition Examination Survey (NHANES): https://wwwn.cdc.gov/nchs/nhanes/Default.aspx
  * National Comorbidity Survey (NCS): https://www.icpsr.umich.edu/web/ICPSR/studies/06693
  * NYC HANES: https://www1.nyc.gov/site/doh/data/data-sets/nyc-hanes-info.page
  * NYC Youth Risk Behavior Study: https://www1.nyc.gov/site/doh/data/data-sets/nyc-youth-risk-behavior-survey.page

